---
title: "DarterTutorial"
author: "Emily Vanaselja"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This package is used to analyze Darter data that has been collected in the southeastern United States. 

How to install:
```{r}

devtools::install_github("evanaselja/darter.pkg")

library(dplyr)
library(ggplot2)
```

In this tutorial we are using gutselli meristics data collected from Tennessee. 

```{r}
#install.packages("googlesheets4")
library(googlesheets4)
googlesheets4::gs4_deauth()
darter_data <- read_sheet("https://docs.google.com/spreadsheets/d/14Xv_bXq3fkIV_zHmRY6dwJZF735Ba9p5YJLQ2wDR0FA/edit?usp=sharing")
```

This package requires ggplot2 and dplyr to be downloaded and added to the directory.

The package contains six functions to evaluate the data:

1. bar_graph Creates a bar graph from any data that is put into the function. The x variable must be numeric. Simply makes a bar graph using the data instead of putting everything into ggplot2.

```{r}

#bar_graph(data, x)

bar_graph(darter_data, "SL")

```

2. clean_data Cleans the data of any NAs, deleting entire row where the NAs are. Cleaning out the NAs is vital in summarizing and quantifying data. 

```{r}

#clean_data(data)

clean_data(darter_data)

```

3. filter_sexes Filters out any sex in a data set. Input the title of the column for sex in for s_vars and the sex you want to eliminate for s.

```{r}

#filter_sexes(data, s_vars, s)

filter_sexes(darter_data, Sex, "m")

```

4. linear_model Creates a linear model. Creates a linear model of the data and the x and y values.

```{r}

#linear_model(df, x, y)

linear_model(darter_data, "SL", "LL")

```

5. plot_data Plots a jitter plot using any data that you put into the function. The y variable must be numeric.

```{r}

#plot_data(data, x, y)

plot_data(darter_data, "SL", "LL")

```

6. summarize_value Summarizes the values of the mean of any variable of the data set, and groups it by any variable.

```{r}

#summarize_value(data, vars, s_var)

summarize_value(darter_data, SL, Sex)

```
